---
title: "ESM 244 HW 3"
author: "Sara Orofino"
date: "2/28/2019"
output: 
  html_document:
  css: custom.css
  toc: true
---
#{.tabset}

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(tseries)
library(forecast)
library(sf)
library(tmap)
library(leaflet)
library(ggrepel)
library(ggspatial)
library(RColorBrewer)
library(raster)
library(effsize)
library(naniar)
```

##Task 1 

Text

##Task 2

###Truckee River Flows

a. Create timeseries data and visualize decomposed data:  

```{r, message=FALSE}

truckee <- read_csv("truckee_flow.csv")

# Convert to ts data
truckee_ts <- ts(truckee$mean_flow, frequency = 12, start = c(2000,1))

#Visualize it:
#plot(truckee_ts)

# Decompose ts data:
truckee_dc <- decompose(truckee_ts)

#Visualize it:
plot(truckee_dc)


```

Explain Timeseries:  

b. Holt-Winters and ARIMA for forecasting:  

```{r}

########### Holt-winters:

truckee_hw <- HoltWinters(truckee_ts)
#plot(truckee_hw)

# Forecast:
truckee_forecast_hw <- forecast(truckee_hw, h = 60)
#plot(truckee_forecast_hw)


########### ARIMA:

# Find pdq with auto arima:
truckee_pdq <- auto.arima(truckee_ts)
#truckee_pdq

# Non-seasonal (2,1,1)
# Seasonal (0,0,2)

# Fit ARIMA with pdq inputs:
truckee_arima <- arima(truckee_ts, order = c(2,1,1), seasonal = list(order = c(0,0,2)))

# Forecast ARIMA:
truckee_forecast_arima <- forecast(truckee_arima, h = 72)
#plot(truckee_forecast_arima)

# Graph of HW: 
plot(truckee_forecast_hw,
     ylab= "Truckee River Flows (cubic feet per second)", 
     xlab = "Months\n(Month 0 = October 2016)") 
```


c. Model Residuals:    

```{r}

# Residuals for HW model:
par(mfrow = c(1,2))
hist(truckee_forecast_hw$residuals) 
qqnorm(truckee_forecast_hw$residuals)


```



##Task 3
###Mapping California National Parks  

```{r}
# Read in CA counties shapefile and set crs to WGS84:

ca_counties <- read_sf(".", layer = "california_county_shape_file")
st_crs(ca_counties) = 4326 


# Read in NPS data and set crs to WGS84:
ca_nps <- read_sf(".", layer = "nps_boundary") %>% 
  filter(STATE == "CA", UNIT_TYPE == "National Park") %>% 
  dplyr::select(UNIT_NAME)
st_crs(ca_nps) = 4326


# CA National Parks graph:
ggplot(ca_counties) +
  geom_sf(fill = "NA",
          color  = "grey30",  
          size =  0.1) +
  geom_sf(data = ca_nps, aes(fill= UNIT_NAME)) 
```


##Task 4
###Lizards in the Northern Chihuahuan Desert 


a. For all lizards trapped at site ‘CALI’, do weights of male and female adult lizards differ
significantly?  

```{r warning=FALSE, message=FALSE}

# Read in data set filter for adult male and female lizards at site 'CALI' (use for both tests)
lizard <- read_csv("lizard_clean.csv") %>% 
  filter(site == "CALI") %>% 
  filter(sex == "M" | sex == "F") %>% 
  replace_with_na_all(condition = ~.x == ".")

# Make weights numeric 
lizard$weight <- as.numeric(lizard$weight)

# Subset data for just male only and female only to visualize distributions 

lizard_male <- lizard %>%
  filter(sex == "M", weight != "NA") 

lizard_female <- lizard %>%
  filter(sex == "F", weight != "NA") 


# Visualize distributions with histograms  

#ggplot(lizard_male, aes(x=weight)) +
  #geom_histogram()

#ggplot(lizard_female, aes(x=weight)) +
  #geom_histogram()

# Both are skewed but sample sizes are over 30 so CLT says means will be evenly distributed 

########## Hypothesis Testing ###############
# Run F-test for equal variances
# HO: Variances are equal (ratio of variances = 1)
# HA: Variances are not equal (ratio of variances ≠ 1)
lizard_ftest <- var.test(weight ~ sex, data = lizard)

# Results - retain the null (p-value = 0.2938) variances are equal 

# Run an upaired t-test to determine if weights are significantly different between males and females: 
# H0: Mean weights are equal    
# HA: Mean weights are not equal  
lizard_ttest <- t.test(weight ~ sex, data = lizard, var.equal = TRUE)

# Results - Mean weights are not significantly different t(130) = 0.79647; p-value = 0.4272 

# Actual mean weights 
male_mean <- mean(lizard_male$weight) # 4.965
male_sd <- sd(lizard_male$weight) #5.679
female_mean <- mean(lizard_female$weight) # 5.826
female_sd <- sd(lizard_female$weight) #6.496

# Difference in mean weights:
mean_diff <- female_mean - male_mean
#0.862

#Effect size Cohen's D:
lizard_eff <- cohen.d(weight ~ sex, data = lizard)
```

Mean adult male lizard body mass (4.97 g ± 5.68 g, n = 57) and female lizard body mass (5.83 g ± 6.50 g, n = 75) did not differ significantly [t(`r lizard_ttest$parameter`) = `r round(lizard_ttest$statistic, 2)`, *p* = `r round(lizard_ttest$p.value, 3)`, $\alpha$ = 0.05]. The actual difference in mean weight between female and male lizards is 0.862 grams with a negligible effect size (Cohen's D = -0.14). 

b. For lizards trapped at the ‘CALI’ site, is there a significant difference in the proportion of
adult male and female lizards with broken tails?    

```{r, warning=FALSE, message=FALSE}

# Chi square to test for differences in proportions 
#H0 - there is no difference in proportion of males and females with broken tails
#HA - there IS a difference in proportion of males and females with broken tails 

lizard_chi <- lizard %>% 
  dplyr::select(sex, tail) %>% 
  filter(tail != "NA") %>% 
  count(sex, tail) %>% 
  spread(tail, n) %>% 
  dplyr::select(-sex)
  
row.names(lizard_chi) <- c("F", "M")

#Perform chi-square 

lizard_x2test <- chisq.test(lizard_chi) 

# NOT a significant difference in proportions X2(1) = 0.30857, p-value = 0.5786

# Proportions for reporting:
# Total male - 56
# Total female - 77
male_broken <- 10/56 #0.179
female_broken <- 18/77 #0.234


```

There is not a significant difference in the proportion of adult male (0.179) and female (0.234) lizards with broken tails [$x^2$ = (`r lizard_x2test$parameter`) = `r round(lizard_x2test$statistic, 2)`, *p* = `r round(lizard_x2test$p.value, 3)`].  

